---
title: "Gridded population of China"
author: "chinaPleth"
date: "April 4th, 2016"
output: 
  html_fragment:
    self_contained: false
    toc: no
---



```{r configuration, message=FALSE, warning=FALSE, include=FALSE}
#opts_knit$set(upload.fun = function(file){library(RWordPress);uploadFile(file)$url;})
```



## Introduction
### Objective
In an earlier [post](http://chinapleth.io/2015/10/24/map-of-china-urbanisation-per-province-with-largest-cities-in-english/) we mapped the urbanization rate of China at province level. In this post we will go futher by visualizing where Chinese people are living using a gridded population map.

We will use the NASA dataset (Population Count Grid, v3 (1990, 1995, 2000)) which consists of estimates of human population for the years 1990, 1995, and 2000 by 2.5 arc-minute grid cells and associated data sets dated circa 2000.

The population count grids contain estimates of the number of persons per grid cell.

### Overall Process :  
1. Get and clean gridded population data as well as grid data




### Required libraries
We need several R packages for this study.

```{r loadLibraries, message=FALSE, warning=FALSE}
library(maptools)
library(raster)
library(gtools)
library(ggplot2)
library(rasterVis)
#library(lubridate)
# Set-up locale for date format processing
Sys.setlocale("LC_ALL","C")
#library(dplyr); ; library(reshape2)
# check if we are in the right working directory
if(gsub("(.*)\\/", "", getwd()) != "Rmd") {setwd("./Rmd")}
```


## Get and clean population and grid data

### Download files
The data is available for free to download on SEDAC website [http://sedac.ciesin.columbia.edu/data/set/gpw-v3-population-count/data-download](http://sedac.ciesin.columbia.edu/data/set/gpw-v3-population-count/data-download)

There are 3 formats available, *.bil, [ *.ascii or *.grid](https://en.wikipedia.org/wiki/Esri_grid)
We will use directly the ASCII version which can be read with `readAsciiGrid()` from the `maptool` package.

We select data for China only and download 3 datasets :  

- Population count grid
- Population density grid
- Land / geographic unit area grid

Each data set is unzipped in a specific folder
```{r configureFileSource}
# p00g	population counts in 2000, unadjusted
popGriCountFile <- "../../data/chinaPlethData/ext/gridded-Population-of-China/chncount-ascii/chnp00g.asc"
# ds00g	population densities in 2000, unadjusted, persons per square km
densGriCountFile <- "../../data/chinaPlethData/ext/gridded-Population-of-China/chndens-ascii/chnds00g.asc"
# areag	land area in square kilometers
areaGriCountFile <- "../../data/chinaPlethData/ext/gridded-Population-of-China/chnarea-ascii/chnareag.asc"
# madmg	mean geographic unit area in square kilometers. These ata are useful for evaluating 	the effective resolution of the input data (the size of the geographic unit area).
admGriCountFile <- "../../data/chinaPlethData/ext/gridded-Population-of-China/chnarea-ascii/chnmadmg.asc"
```



### Load data

```{r loadGridData}
#x <- readAsciiGrid(fname = popGriCountFile)

pop <- raster(popGriCountFile)
fromDisk(pop)

dens <- raster(densGriCountFile)
fromDisk(dens)

area <- raster(areaGriCountFile)
fromDisk(area)

adm <- raster(admGriCountFile)
fromDisk(adm)

```


## Visualise gridded data
```{r visualizeData}
plot(pop)
plot(dens)
plot(area)
plot(adm)



boundingBox <- c(115, 125, 25, 35)
pop.crop <- crop(pop, boundingBox)
plot(pop.crop)

```


## Map of China
```{r}
ChinaPolygonsLevel1 <- readShapeSpatial("../../data/chinaPlethData/ext/GAA/CHN_adm1.shp")

## Fix English names, simplify
ChinaPolygonsLevel1@data$NAME_1 <- as.character(ChinaPolygonsLevel1@data$NAME_1)
ChinaPolygonsLevel1@data[grep("Xinjiang Uygur", ChinaPolygonsLevel1@data$NAME_1),]$NAME_1 <- "Xinjiang"
ChinaPolygonsLevel1@data[grep("Nei Mongol", ChinaPolygonsLevel1@data$NAME_1),]$NAME_1 <- "Nei Menggu"
ChinaPolygonsLevel1@data[grep("Ningxia Hui", ChinaPolygonsLevel1@data$NAME_1),]$NAME_1 <- "Ningxia"
ChinaPolygonsLevel1@data$NAME_1 <- as.factor(ChinaPolygonsLevel1@data$NAME_1)

# Use level1 as index & Province name as id
ChinaLevel1Data <- ChinaPolygonsLevel1@data
ChinaLevel1Data$id <- ChinaLevel1Data$NAME_1

# Fortify the data (polygon map as dataframe) using english names
ChinaLevel1dF <- fortify(ChinaPolygonsLevel1, region = "NAME_1")

## Merge polygons and associated data in one data frame by id (name of the province in chinese)
ChinaLevel1 <- merge(ChinaLevel1dF, ChinaLevel1Data, by = "id")
rm(ChinaPolygonsLevel1, ChinaLevel1dF, ChinaLevel1Data)


## convert the raster to points (to plot with ggplot)
raster.points <- rasterToPoints(pop)
raster.points <- data.frame(raster.points)
colnames(raster.points) <-c('x','y','population')

## find x% of China
total.pop <- sum(raster.points$population)
n.unit <- nrow(raster.points)
av.unit.pop <- total.pop / n.unit

sum(raster.points[raster.points$population > av.unit.pop,]$population)

mp <- ggplot(data=raster.points[raster.points$population > 1,], aes(y=y, x=x)) 
mp <- mp + geom_raster(aes(fill=log(population)))
mp

## Create the ggplot using standard approach
## group is necessary to draw in correct order, try without to understand the problem
g <- ggplot(ChinaLevel1, aes(x = long, y = lat, fill = ENGTYPE_1, group = group))

## GGPLOT
g <- ggplot()
## projected shadow
g <- g+ geom_polygon(data = ChinaLevel1, aes(x = long + 0.7, y = lat - 0.5, group = group), color = "grey50", size = 0.2, fill = "grey50")

## Add Raster
g <- g + geom_raster(data=raster.points[raster.points$population > 1,], aes(y=y, x=x, fill=population))

## Province boundaries
g <- g + geom_point(data = ChinaLevel1, aes(x = long, y = lat), color = "white", size = 0.2)

## to keep correct ratio in the projection
g <- g + coord_equal()
g <- g + labs(title = "China - level 1 subdivisions")
print(g)
```



## References
### Similar work
http://metrocosm.com/world-population-split-in-half-map/

Center for International Earth Science Information Network - CIESIN - Columbia University, United Nations Food and Agriculture Programme - FAO, and Centro Internacional de Agricultura Tropical - CIAT. 2005. Gridded Population of the World, Version 3 (GPWv3): Population Count Grid. Palisades, NY: NASA Socioeconomic Data and Applications Center (SEDAC). http://dx.doi.org/10.7927/H4639MPP. Accessed DAY MONTH YEAR. 

### How to
https://pakillo.github.io/R-GIS-tutorial/#raster
http://zevross.com/blog/2015/03/30/map-and-analyze-raster-data-in-r/
https://jeffreybreen.wordpress.com/tag/raster/
https://rstudio-pubs-static.s3.amazonaws.com/68747_1ce236d66dcd4d1e875ca34a51344530.html
http://nrelscience.org/2013/05/30/this-is-how-i-did-it-mapping-in-r-with-ggplot2/