---
title: "Getting and cleaning data, example of Chinese airports - Part 5/5"
author: "chinaPleth"
date: "January 1, 2015"
output: 
  html_document: 
    toc: yes
---

## Introduction
### Objective  
One of the big problem for anybody interested in China and data science is the availability of data sets. There are limited free ressources available and they are often incomplete or innacurate. Getting data and especially cleaning data becomes one of the biggest pain of data science applied to China.

The objective of this group of post is to illustrate the problem and associated process on a specific example: plot a map of the airports of mainland China.

In this fifth part, we will fix, clean data for the complement data set generated in part [part 4](http://chinapleth.io/2016/01/01/getting-and-cleaning-data-example-of-chinese-airports-part-45/)


### Dependencies
We will need few standard packages as loaded bellow.

```{r libLoading, message=FALSE}
library(maptools); library(reshape2)
library(grid); library(dplyr)
library(ggplot2); library(gmaps)
library(ggmap); library(jsonlite)
library(Amelia)
```

### Loading part 4 objects

```{r loadPart1, echo=TRUE,eval=TRUE}
load(file = "./chinese-airports-part-4.Rda")
```



## Finding missing data for remaining airports (`airports_2_chec` dataset)
### What's our plan ?
```{r planA2C}
missmap(airports_2_check, main = "Missingness map for remaining airports")
```

This set of aiports have a lot of data missing, we will use the wikipedia dataset to complete it as much as possible.

### Finding missing `ICAO` and missing `IATA.FAA`

Let's first complete missing ICAO for known IATA.FAA code and missing IATA.FAA for known ICAO codes.

```{r A2CMissingICAO}
## Search pattern for airport with IATA.FAA codes and no ICAO code
search <- !is.na(airports_2_check$IATA.FAA) & is.na(airports_2_check$ICAO)
missing.icao.with.iata <- airports_2_check[search,]$IATA.FAA
## Look in wikipedia data set
found.icao <- wikiAirports[wikiAirports$IATA.FAA %in% missing.icao.with.iata,]
found.icao <- dplyr::select(found.icao, IATA.FAA, ICAO)
airports_2_check[airports_2_check$IATA.FAA %in% found.icao$IATA.FAA,]$ICAO <- found.icao$ICAO

## Search pattern for airport with ICAO codes and no IATA.FAA code
search <- is.na(airports_2_check$IATA.FAA) & !is.na(airports_2_check$ICAO)
missing.iata.with.icao <- airports_2_check[search,]$ICAO
## Look in wikipedia data set
found.iata <- wikiAirports[wikiAirports$ICAO %in% missing.iata.with.icao,]
found.iata <- dplyr::select(found.iata, IATA.FAA, ICAO)
airports_2_check[airports_2_check$ICAO %in% found.iata$ICAO,]$IATA.FAA <- found.iata$IATA.FAA
missmap(airports_2_check)
```


### Removing military airports
We find out that most of airports with missing ICAO and IATA.FAA are in fact military bases. Let's remove these.

```{r A2CremovingMilitary}
# search pattern
search <- is.na(airports_2_check$IATA.FAA) & is.na(airports_2_check$ICAO) & grepl("AIRFIELD|TEST RANGE|NAVAL BASE|REEF|AIR BASE", toupper(airports_2_check$name))
# check names
airports_2_check[search,]$name
# remove these lines
airports_2_check <- airports_2_check[!search,]
```


### Removing airports without ICAO or IATA
We find out that remaining airports are either already in IATA data set (such as Lijiang airport) or closed, we decide to remove all airports if IATA.FAAA and ICAO code are missing

```{r A2CcleanCodes}
# search pattern
search <- is.na(airports_2_check$IATA.FAA) & is.na(airports_2_check$ICAO)
# check names
airports_2_check[search,]$name
# remove these lines
airports_2_check <- airports_2_check[!search,]
missmap(airports_2_check)
```


### Removing double with IATA dataframe
There are some airports which are already in the IATA dataset, we chose to remove these.

```{r remove2}
## check if some IATA codes of airports_2check are double
checkIATA <- as.data.frame(table(airports_2_check$IATA.FAA))
names(checkIATA) <- c("IATA.FAA", "freq")
checkIATA[checkIATA$freq > 1,]$IATA.FAA

## check if some IATA codes of airports_2check are in IATA dataframe
checkIATA <- as.data.frame(table(c(IATA$IATA.FAA, airports_2_check$IATA.FAA)))
names(checkIATA) <- c("IATA.FAA", "freq")
a <- checkIATA[checkIATA$freq > 1,]$IATA.FAA
a1 <- dplyr::select(IATA[IATA$IATA.FAA %in% a,], IATA.FAA, ICAO, name, latitude.deg, longitude.deg)
a1$source <- "IATA"
a2 <- dplyr::select(airports_2_check[airports_2_check$IATA.FAA %in% a,], IATA.FAA, ICAO, name, latitude.deg, longitude.deg)
a2$source <- "airports_2_check"
rbind(a1,a2)
airports_2_check <- airports_2_check[!(airports_2_check$IATA.FAA %in% a),]
```

```{r remove3}
## check if some ICAO codes of airports_2check are double
checkICAO <- as.data.frame(table(airports_2_check$ICAO))
names(checkICAO) <- c("ICAO", "freq")
a <- checkICAO[checkICAO$freq > 1,]$ICAO
a1 <- dplyr::select(airports_2_check[airports_2_check$ICAO %in% a,], IATA.FAA, ICAO, name, latitude.deg, longitude.deg)
a2 <- dplyr::select(wikiAirports[wikiAirports$ICAO %in% a,], IATA.FAA, ICAO, name)
a2
#a2$latitude.deg <- 0
#a2$longitude.deg <- 0
#rbind(a1,a2)
## remove these airports, it's difficult to conclude which data is correct
#airports_2_check <- airports_2_check[!(airports_2_check$IATA.FAA %in% a),]
```

```{r remove4}
## check if some IATA codes of airports_2check are in IATA dataframe
checkICAO <- as.data.frame(table(c(IATA$ICAO, airports_2_check$ICAO)))
names(checkICAO) <- c("ICAO", "freq")
b <- checkICAO[checkICAO$freq > 1,]$ICAO
b1 <- dplyr::select(IATA[IATA$ICAO %in% b,], IATA.FAA, ICAO, name, latitude.deg, longitude.deg)
b1$source <- "IATA"
b2 <- dplyr::select(airports_2_check[airports_2_check$ICAO %in% b,], IATA.FAA, ICAO, name, latitude.deg, longitude.deg)
b2$source <- "airports_2_check"
rbind(b1,b2)
airports_2_check <- airports_2_check[!(airports_2_check$IATA.FAA %in% a),]

```

### Fixing missing data with wikipedia
We use wikipedia data frame from previous section to fix missing information. We can fix few missing ICAO codes but no missing IATA.

```{r AC2fixmissing}
## search AO missing ICAO with IATA in wikiairport
search <- wikiAirports$IATA.FAA %in% airports_2_check[is.na(airports_2_check$ICAO),]$IATA.FAA
## replace their ICAO code
search2 <- airports_2_check$IATA.FAA %in%  wikiAirports[search,]$IATA.FAA 
airports_2_check[search2,]$IATA.FAA
airports_2_check[search2,]$ICAO <- wikiAirports[search,]$ICAO
missmap(airports_2_check)

## search AO missing IATA with ICAO in wikiairport
sum(wikiAirports$ICAO %in% airports_2_check[is.na(airports_2_check$IATA.FAA),]$ICAO)
```


```{r AC2fixmissingO}
## check wikipedia data for airport with IATA Code
A1 <- merge(airports_2_check[!is.na(airports_2_check$IATA.FAA),], wikiAirports, by = "IATA.FAA", all.x = FALSE)
## check ICAO code difference
A1$distICAO <- stringdist(A1$ICAO.x, A1$ICAO.y)
A1$distName <- stringdist(A1$name.x, A1$name.y)
A1$distLink <- stringdist(A1$airport.wikipedia.link.x, A1$airport.wikipedia.link.x)



## keep wikipedia names for airport with same ICAO but difference in name.x and name.y
search <- !is.na(A1$distName) & A1$distName > 0 & !is.na(A1$distICAO) & A1$distICAO == 0
dplyr::select(A1[search,], name.x, name.y, IATA.FAA)
A1[search,]$name.x <- A1[search,]$name.y
A1$distName <- stringdist(A1$name.x, A1$name.y)

## keep ICAO from wikipedia for airport with same names but difference in ICAO.x and ICAO.y
search <- !is.na(A1$distName) & A1$distName == 0 & !is.na(A1$distICAO) & A1$distICAO > 0
dplyr::select(A1[search,], name.x, name.y, IATA.FAA, ICAO.x, ICAO.y)
A1[search,]$ICAO.x <- A1[search,]$ICAO.y
A1$distICAO <- stringdist(A1$ICAO.x, A1$ICAO.y)

## checkout mismatch in ICAO.x and ICAO.y 
search <- !is.na(A1$distICAO) & A1$distICAO > 0
a <- dplyr::select(A1[search,], name.x, name.y, IATA.FAA, ICAO.x, ICAO.y)

for(i in 1:nrow(a)) {
        print(IATA[IATA$ICAO == a[i,]$ICAO.x & !is.na(IATA$ICAO),]$latitude.deg)
}
a <- dplyr::select(A1, name.x, name.y, IATA.FAA, ICAO.x, ICAO.y, distName, distICAO, distLink, longitude.deg, latitude.deg)



```




checkairports_2_check <- as.data.frame(table(airports_2_check[!is.na(airports_2_check$IATA.FAA),]$IATA.FAA))
names(checkairports_2_check) <- c("IATA.FAA", "freq")
checkairports_2_check[checkairports_2_check$freq > 1,]$IATA.FAA
checkairports_2_check <- as.data.frame(table(airports_2_check[!is.na(airports_2_check$ICAO),]$ICAO))
names(checkairports_2_check) <- c("ICAO", "freq")
checkairports_2_check[checkairports_2_check$freq > 1,]$ICAO

```

