---
title: "chinaPleth/aqi dataset Codebook"
author: "chinaPleth"
date: "March 22, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
load("../Rmd/aqi-1.Rda")
```


## Introduction and objectives

This document details the content of the chinaPleth/aqi dataset.
Read more on how to obtain this data [here](http://chinapleth.io/2016/01/09/about-polution-particle-matters-pm2-5-beijing-vs-shanghai-and-other-cities-part-1/)

## About aqi dataset

This dataset contains hourly pm2.5 (particle matters) for five Chinese cities (Beijing, Chengdu, Guangzhou, Shanghai, Shenyang) and Paris.

## Structure of the dataset

```{r}
str(aqi)
```

## Comments on variables

Variable | Values | Comments
---------|--------|---------
country | `r unique(aqi$country)` | in English
city | `r unique(aqi$city)` | in English
date.time | POSIXlt date and time | format yyy-mm-dd hh:mm:ss to match with hourly readings
date | chr  | in yyyy-mm-dd format
year | num | yyyy format
month | num | 1 to 12
day | num | 1 to 31
hour | num | 0 to 23
measurement.available | chr | `r unique(aqi$measurement.available)`
pm2.5 | num | particle matters (pm2.5) readings in micro gram per cubic meter
aqi | num | aqi readings according to american standard



